Plots

### 2025-10-24 src/plots.Rmd
Use and rerun for stateless gpt5
Add conf intervals, working on last respones to reviewer2

---
title: "Plots"
date: "`r format(Sys.time(), '%B %d, %Y')`"
author: "Richard Wen <richard.wen@unityhealth.to>"
---

@@

```{r, echo=FALSE, eval=FALSE}
#install.packages("openVA")
#install.packages("tidyverse")
#install.packages("dplyr")
#install.packages("ggrepel")
#install.packages("patchwork")
#install.packages("glue")
#install.packages("kableExtra)
```

@@

```{r}
library(tidyverse)
library(dplyr)
library(ggrepel)
library(ggpattern)
library(patchwork)
library(glue)
library(kableExtra)
```

@@

    # Calc PCCC
    out <- (C - (k/N)) / (1 - (k/N))

@@

@@

# Plot performance of models using grouped bars
plot_perf_bar <- function(
        df_long,
        f = calc_pccc,
        group = "cod_name",
        color = "model_name",
        text_scale = 1,
        title = NULL,
        title_size = 10,
        x_title = "Partial Chance Corrected Concordance (PCCC)",
        y_title = NULL,
        x_hpos = -0.23 * text_scale,
        x_limits = c(0, 1),
        x_breaks = seq(0, 1, by = 0.1),
        x_expand = expansion(add = ((x_hpos * -1) + 0.02) * text_scale),
        x_shapes = 21:26,
        x_title_size = 9 * text_scale,
        y_title_size = 9 * text_scale,
        x_text_size = 8 * text_scale,
        x_color_size = 3 * text_scale,
        y_text_size = 9 * text_scale,
        legend_position = "bottom",
        legend_title = element_blank(),
        legend_margin = margin(c(-8, -2, -2, -50)),
        legend_key_size = 0.5,
        ...
    ) {

@@

@@

    # Reverse sort groups by highest-performing bar
    group_levels_order <- df_pos %>%
        group_by(y) %>%
        summarize(max_x = max(x)) %>%
        arrange(max_x) %>%
        pull(y)

    df_pos <- df_pos %>%
        mutate(
            y = factor(y, levels = group_levels_order),
            group_center = as.numeric(y)
        )

    # Bar width and offset
    group_width <- 0.8
    max_items <- max(df_pos$n_colors)
    bar_width <- group_width / max_items

    df_pos <- df_pos %>%
        mutate(
            offset = (rank_in_group - (n_colors + 1)/2) * bar_width,
            x_pos = group_center + offset,
            label_text = paste0(color, ": ", format(round(x, 2), nsmall = 2)),
            label_x = x_pos
        )

@@

    # Plot metric for each model
    out <- ggplot(df_pos, aes(x = x_pos, y = x, fill = color)) +
        geom_col(
            width = bar_width * 0.9,
            position = "identity",
            color = df_pos$outline_color,
            alpha = df_pos$alpha_val
        ) +
        geom_text(
            aes(
                x = label_x,
                label = label_text,
                fontface = fontface
            ),
            y = x_hpos,
            hjust = 0,
            vjust = 0.5,
            size = x_color_size
        ) +
        coord_flip() +
        scale_fill_manual(
            values = default_colors
        ) +
        scale_x_continuous(
            breaks = seq_along(levels(df_pos$y)),
            labels = levels(df_pos$y),
            expand = c(0, 0)
        ) +
        scale_y_continuous(
            expand = x_expand,
            breaks = x_breaks,
            limits = x_limits
        ) +
        labs(
            title = title,
            x = y_title,
            y = x_title
        ) +
        theme_minimal() +
        theme(
            legend.position = legend_position,
            legend.title = legend_title,
            legend.margin = legend_margin,
            legend.key.size = unit(legend_key_size, "lines"),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            plot.title = element_text(size = title_size),
            axis.title.x = element_text(size = x_title_size),
            axis.title.y = element_text(size = y_title_size),
            axis.text.x = element_text(size = x_text_size),
            axis.text.y = element_text(size = y_text_size)
        )
    return(out)
}

@@

# Plot perf for pccc by age groups
fig_pccc <- plot_perf_bar(
    df_long %>%
        mutate(
            age_label = case_when(
                age_name == "Adult"
                    ~ paste0(age_name, "\n12+ Years\n(n=", n_adult, ")"),
                age_name == "Child"
                    ~ paste0(age_name, "\n28 Days to 11 Years\n(n=", n_child, ")"),
                age_name == "Neonate"
                    ~ paste0(age_name, "\n<28 Days\n(n=", n_neo, ")"),
                .default = age_name
            ) %>%
                factor(., levels = unique(.) %>% sort %>% rev)
        ),
    calc_pccc,
    group = "age_label",
    x_hpos = -0.24,
    x_expand = expansion(add = 0.25),
    text_scale = 0.95
)

@@

# Plot adult perf for pccc by cod
fig_pccc_cod_adult <- plot_perf_bar(
    df_long %>% filter(age == "adult"),
    calc_pccc,
    group = "cod_name",
    title = paste0("Adult Causes of Death, 12+ Years (n=", n_adult, ")"),
    text_scale = 0.7,
    x_hpos = -0.15,
    x_expand = expansion(add = 0.15)
)

@@

# Plot child perf for pccc by cod
fig_pccc_cod_child <- plot_perf_bar(
    df_long %>% filter(age == "child"),
    calc_pccc,
    group = "cod_name",
    title = paste0("Child Causes of Death, 28 Days to 11 Years (n=", n_child, ")"),
    text_scale = 0.92,
    x_hpos = -0.24,
    x_expand = expansion(add = 0.25)
)

@@

# Plot neo perf for pccc by cod
fig_pccc_cod_neo <- plot_perf_bar(
    df_long %>% filter(age == "neo"),
    calc_pccc,
    group = "cod_name",
    title = paste0("Neonatal Causes of Death, <28 Days (n=", n_neo, ")"),
    x_title = "Partial Chance Corrected Concordance (PCCC)",
    text_scale = 0.92,
    x_hpos = -0.24,
    x_expand = expansion(add = 0.25)
)

>>>>

---
title: "Plots"
date: "`r format(Sys.time(), '%B %d, %Y')`"
author: "Richard Wen <rrwen.dev@gmail.com>"
---

@@

```{r, echo=FALSE, eval=FALSE}
#install.packages("openVA")
#install.packages("tidyverse")
#install.packages("dplyr")
#install.packages("ggrepel")
#install.packages("patchwork")
#install.packages("glue")
#install.packages("kableExtra)
#install.packages("boot")
```

@@

```{r}
library(tidyverse)
library(dplyr)
library(ggrepel)
library(ggpattern)
library(patchwork)
library(glue)
library(kableExtra)
library(boot)
```

@@

    # Calc PCCC
    if (!all(TP)) {
        out <- (C - (k/N)) / (1 - (k/N))
    } else { # avoid division by 0 when max pccc
        out <- 1
    }

@@

# Calc bootstrapped cis for a metric
calc_ci <- function(
        x,
        f = calc_pccc,
        n_boot = 1000,
        conf_level = 0.95,
        col_actual = "cod",
        col_pred = "model_cghr10",
        ...
    ) {

    # Function to compute metric on a bootstrap sample
    f_boot <- function(df, i) {
        f(df[[col_actual]][i], df[[col_pred]][i], ...)
    }

    # Run bootstrap using the 'boot' package
    boot_res <- boot(data = x, statistic = f_boot, R = n_boot)

    # Compute percentile-based confidence interval
    ci <- boot.ci(boot_res, type = "perc", conf = conf_level)$percent[4:5]

    # Return named vector with metric and CI
    out <- c(ci_mean = mean(boot_res$t), ci_low = ci[1], ci_hi = ci[2])
    return(out)
}

@@

# Plot performance of models using grouped bars
plot_perf_bar <- function(
        df_long,
        f = calc_pccc,
        group = "cod_name",
        color = "model_name",
        text_scale = 1,
        title = NULL,
        title_size = 10,
        x_title = "Partial Chance Corrected Concordance (PCCC)",
        y_title = NULL,
        x_hpos = -0.47 * text_scale,
        x_limits = c(0, 1),
        x_breaks = seq(0, 1, by = 0.1),
        x_expand = expansion(add = ((x_hpos * -1) + 0.03)),
        x_shapes = 21:26,
        x_title_size = 8 * text_scale,
        y_title_size = 8 * text_scale,
        x_text_size = 7 * text_scale,
        x_color_size = 2.5 * text_scale,
        y_text_size = 8 * text_scale,
        legend_position = "bottom",
        legend_title = element_blank(),
        legend_margin = margin(c(-8, -2, -2, -50)),
        legend_key_size = 0.5,
        legend_text_size = 7 * text_scale,
        legend_spacing_x = unit(58, "pt"),
        n_boot = 1000,
        error_bar = T,
        error_bar_width = 0.05,
        error_bar_size = 0.3,
        error_bar_label = "95% Confidence Interval (Bootstrap, 1000 Resamples)"
    ) {

@@

    # Calc metric
    if (identical(f, calc_pccc)) {
        df_plot <- df_plot %>%
            group_map(~{

                # Calc metric
                out <- c(
                    metric = f(.x$cod, .x$model_cghr10, N = unique(.x$ncod)[1])
                )

                # Include ci if error bar
                if (error_bar) {
                    out <- c(.y, out, calc_ci(.x, n_boot = n_boot))
                }
                return(out)

            }, .keep = T) %>%
            bind_rows

    } else {
        df_plot <- df_plot %>%
            group_map(~{

                # Calc metric
                out <- c(
                    metric = f(.x$cod, .x$model_cghr10)
                )

                # Include ci if error bar
                if (error_bar) {
                    out <- c(.y, out, calc_ci(.x, n_boot = n_boot))
                }
                return(out)

            }, .keep = T) %>%
            bind_rows
    }

@@

    # Reverse sort groups by highest-performing bar
    group_levels_order <- df_pos %>%
        group_by(y) %>%
        summarize(max_x = max(x)) %>%
        arrange(max_x) %>%
        pull(y)
    df_pos <- df_pos %>%
        mutate(
            y = factor(y, levels = group_levels_order),
            group_center = as.numeric(y)
        )

    # Bar width and offset
    group_width <- 0.8
    max_items <- max(df_pos$n_colors)
    bar_width <- group_width / max_items
    df_pos <- df_pos %>%
        mutate(
            offset = (rank_in_group - (n_colors + 1)/2) * bar_width,
            x_pos = group_center + offset,
            label_text = paste0(color, ": ", format(round(x, 2), nsmall = 2)),
            label_x = x_pos
        )

    # Add ci to model label
    if (error_bar) {
        df_pos <- df_pos %>%
            mutate(
                label_text = paste0(label_text, " (", round(ci_low, 2), ",", round(ci_hi, 2), ")")
            )
    }

@@

    # Plot metric for each model
    df_pos$ci_label <- error_bar_label # dummy label for legend
    out <- ggplot(df_pos, aes(x = x_pos, y = x, fill = color)) +
        geom_col(
            width = bar_width * 0.9,
            position = "identity",
            color = df_pos$outline_color,
            alpha = df_pos$alpha_val
        ) +
        geom_text(
            aes(
                x = label_x,
                label = label_text,
                fontface = fontface
            ),
            y = x_hpos,
            hjust = 0,
            vjust = 0.5,
            size = x_color_size
        ) +
        coord_flip() +
        scale_fill_manual(
            values = default_colors
        ) +
        scale_x_continuous(
            breaks = seq_along(levels(df_pos$y)),
            labels = levels(df_pos$y),
            expand = c(0, 0)
        ) +
        scale_y_continuous(
            expand = x_expand,
            breaks = x_breaks,
            limits = x_limits
        ) +
        labs(
            title = title,
            x = y_title,
            y = x_title
        ) +
        theme_minimal() +
        theme(
            legend.position = legend_position,
            legend.title = legend_title,
            legend.margin = legend_margin,
            legend.key.size = unit(legend_key_size, "lines"),
            legend.text = element_text(size = legend_text_size),
            legend.spacing.x = legend_spacing_x,
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            plot.title = element_text(size = title_size),
            axis.title.x = element_text(size = x_title_size),
            axis.title.y = element_text(size = y_title_size),
            axis.text.x = element_text(size = x_text_size),
            axis.text.y = element_text(size = y_text_size)
        )

    # Show error bars
    if (error_bar) {
        out <- out +
            geom_errorbar(
                aes(ymin = ci_low, ymax = ci_hi, color = ci_label),
                width = error_bar_width,
                size = error_bar_size
            ) +
            scale_color_manual(name = "", values = "black") +
            guides(
                fill = guide_legend(order = 1), # bar legend first
                color = guide_legend(order = 2) # ci legend last
            )
    }
    return(out)
}

@@

# Plot perf for pccc by age groups
fig_pccc <- plot_perf_bar(
    df_long %>%
        mutate(
            age_label = case_when(
                age_name == "Adult"
                    ~ paste0(age_name, "\n12+ Years\n(n=", n_adult, ")"),
                age_name == "Child"
                    ~ paste0(age_name, "\n28 Days to 11 Years\n(n=", n_child, ")"),
                age_name == "Neonate"
                    ~ paste0(age_name, "\n<28 Days\n(n=", n_neo, ")"),
                .default = age_name
            ) %>%
                factor(., levels = unique(.) %>% sort %>% rev)
        ),
    calc_pccc,
    group = "age_label",
    text_scale = 0.95
)

@@

# Plot adult perf for pccc by cod
fig_pccc_cod_adult <- plot_perf_bar(
    df_long %>% filter(age == "adult"),
    calc_pccc,
    group = "cod_name",
    title = paste0("Adult Causes of Death, 12+ Years (n=", n_adult, ")"),
    text_scale = 0.7,
    x_hpos = -0.27
)

@@

# Plot child perf for pccc by cod
fig_pccc_cod_child <- plot_perf_bar(
    df_long %>% filter(age == "child"),
    calc_pccc,
    group = "cod_name",
    title = paste0("Child Causes of Death, 28 Days to 11 Years (n=", n_child, ")"),
    text_scale = 0.92,
    x_hpos = -0.43
)

@@

# Plot neo perf for pccc by cod
fig_pccc_cod_neo <- plot_perf_bar(
    df_long %>% filter(age == "neo"),
    calc_pccc,
    group = "cod_name",
    title = paste0("Neonatal Causes of Death, <28 Days (n=", n_neo, ")"),
    x_title = "Partial Chance Corrected Concordance (PCCC)",
    text_scale = 0.92,
    x_hpos = -0.43
)
